---
title: "Stock portfolio summary"
author: "BScheliga"
date: "19/11/2021"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(quantmod) # to get the stock exchange data
library(tidyverse) # for plotting
library(vroom)
library(lubridate)
library(ggthemes)
library(webr)# for PieDonut()
library(here)

### ggplot theme
theme_set(theme_bw())

```

## Abstract

This is script allow you to assess the performance, dividend yield and the sector and market distribution of your stock portfolio. It retrieves the historical stock records based in the EPIC from [https://www.alphavantage.co/](https://www.alphavantage.co/). The usage of Alpha Vantage requires an API-key, which can be requested here [https://www.alphavantage.co/support/#api-key](https://www.alphavantage.co/support/#api-key) 

### Note:
* change wd to more appropiate
* Add geography category (e.g. Continents or countries maybe)


### Known (potential) issues:
* freshly bought stocks won't show up in the summary until a daily value is available
* for app with data entry tool. the user need to investigate, if stock it traded in GPX or GBP
* Initial I used Yahoo.finace as a data source, but their data was of poor quality. e.g. inconsistent in stock value unit, a timeseries could have GPX and GPB. Hence, I changed to Alpha Vatage. However, they require you to have an API.KEY.
* identical EPIC's at different Stock exchange, will cause a problems in some of the loops where *vec_EPIC* is used
* when calculating the investments selling stock with a lose will generate positive values in the cs_???_Inv columns. However, that money is actually lost :(
* not sure how the plots respond to Stock being completetly sold and later repurchased.

```{r load your portfolio data, include=FALSE}
# this is empty
dir()
# loading the dataset with the trading activities
df_PF_input <- vroom(here("ETF_Share.csv"),col_types = c(EPIC = "f", Name = "f", Buy_Sell = "f", Type = "f", Sector = "f", Market = "f")) # "f" for factor, 

# Formatting the Date column. 
df_PF_input <- mutate(df_PF_input,Date = dmy(Date)) 


# loading the dataset with the dividend history
df_PF_div <- vroom(here("dividend.csv"),col_types = c(EPIC = "f", Name = "f", Type = "f")) # "f" for factor, 

# Formatting the Date column. 
df_PF_div <- mutate(df_PF_div,Issue_Date = dmy(Issue_Date))
df_PF_div <- mutate(df_PF_div,XD_Date = dmy(XD_Date))

```


```{r Prelude stuff (extract EPIC, cumsums etc), include=FALSE}
# here we will extract the EPIC from the .cvs file into a vector to use them in the getSymbol chunk
vec_EPIC <- levels(df_PF_input$EPIC)

# reorder dt_PF_input after the dates
df_PF_input <- df_PF_input[order(df_PF_input$Date),]
# Changing the sold Volumes of a stock to negative values for the cumulative sums
df_PF_input$Volume <- ifelse(df_PF_input$Buy_Sell == "BUY",df_PF_input$Volume, -(df_PF_input$Volume))

# Calculating the cumulative sum
## it is easier (for me) to create a separate df here add the cumulative sum (cs) for the trading costs later to time series. It prevents the cs_Trad_cost "fluctuate" later between the EPIC on the same day as the fill-function for the time series doesn't adjust the cs trading costs. Don't worry, if the explanation doesn't make sense. Just know it is needed. Peace out.
df_cs_Trad_cost <- data.frame(Date = df_PF_input$Date, cs_Trad_cost = df_PF_input$Trad_cost)
df_cs_Trad_cost <- aggregate(df_cs_Trad_cost["cs_Trad_cost"], by=df_cs_Trad_cost["Date"],sum) # We'll need that step, in case more than one trade action was made per day
df_cs_Trad_cost$cs_Trad_cost <- cumsum(df_cs_Trad_cost$cs_Trad_cost)

#### Calculating the cumulative sum for each stock (EPIC) in our portfolio 
df_PF_input <- df_PF_input %>% group_by(EPIC) %>% mutate(cs_Volume = cumsum(Volume))%>% ungroup()

### Calculating overall investment through the booking cost based on average cost per stock and number of stock bought. Selling stocks result in "negative" investment. the overall investment is a cumulative sum

## per indivual stock
df_PF_input <- df_PF_input %>% group_by(EPIC) %>% mutate(cs_Stock_Inv = round(cumsum(Aver_cost*Volume/100),2)) %>% ungroup()

## overall fo the whole portfolio NOTE: Daily values
df_PF_input <- df_PF_input %>% mutate(cs_All_Inv = round(Aver_cost*Volume/100,2))


df_All_Inv <- aggregate(df_PF_input["cs_All_Inv"],by = df_PF_input["Date"], sum )
df_All_Inv$cs_All_Inv <- cumsum(df_All_Inv $cs_All_Inv)


## per type (ETF and Share) NOTE: Daily values
df_PF_input <- df_PF_input %>% group_by(Type) %>% mutate(cs_ETF_Share_Inv = round(Aver_cost*Volume/100,2)) %>% ungroup()

df_ETF_Share_Inv <- aggregate(df_PF_input["cs_ETF_Share_Inv"],by = df_PF_input[c("Date", "Type")], sum )
df_ETF_Share_Inv <- df_ETF_Share_Inv %>% group_by(Type) %>% mutate(cs_ETF_Share_Inv = cumsum(cs_ETF_Share_Inv)) %>% ungroup()


```

```{r Setup API.KEY revolver, include=FALSE}
# If available this section would allow you to cycle through multiple API-keys
# One API can do 5 requests per hour aka request 5 stock timeseries
API_KEY_1 <- "23OWOXACHQ1GPHF3"
API_KEY_2 <- "23OWOXACHQ1GPHF3"
API_KEY_3 <- "23OWOXACHQ1GPHF3"
API_KEY_4 <- "23OWOXACHQ1GPHF3"
API_KEY_5 <- "23OWOXACHQ1GPHF3"

## store all the names of the API-Key object with an 
## this will be used to cycle through API-keys with each request
API_ABDN <- ls(pattern="API_KEY")

# to keep track of how many are left of the day, 5*500=2500 per day
API_counter <- 500

```




```{r getSymbol - Get stock exchange data, include=FALSE}
# we are using the vec_EPIC with our stock EPIC's to retrieve the historical data for each stock using getSymbols()
# Note need the ".L" for London stock exchange. NASDAQ is the default

df_StockExch <- ".L" # I know, not a data frame, yet. However, it is likely that is will be one in a later version to be able pull the hist. data for stocks at different Stock exchange in a portfolio

# Setup 
keycnt=1 # will be used to go through the API's

for(i in 1:length(vec_EPIC)){
  start_time <- Sys.time() # when the loop starts
  getSymbols(paste(vec_EPIC[i],df_StockExch,sep = ""), src="av", api.key=get(API_ABDN[keycnt]), periodicity="daily", output.size="full")
  API_counter <- API_counter-1
  print(get(API_ABDN[keycnt]))
print(paste(API_counter, "requests left on API", sep = " "))
  keycnt = keycnt + 1
  Sys.sleep(12) # wait for 12 secs after each calls to comply with the API limitation. Not more the 5 calls per minute.
    if (keycnt > length(API_ABDN)){
      keycnt = 1
        }
}

```

```{r create timeserie for our stock portfolio, include=FALSE}
#Note: getSymbol returns a xts-object which does not have rownames. But the dates as index. So, we need to use index to access the data

#First we are creating a new larger df with all dates from point of purchase until now
# following this solutions https://stackoverflow.com/questions/50530914/how-to-generate-date-sequence-between-two-dates-in-same-column-in-r
df_ETF_Share_ts <- df_PF_input %>% group_by(EPIC)%>% # for each share, we can potentially delete that later
  complete(Date = seq.Date(min(Date), max(Sys.Date()), by= "day")) %>% # Generates the dates between the Date. Sys.Date for update current date
  fill(EPIC, Name, Type, Market, Sector, cs_Volume, cs_Stock_Inv) %>% ungroup()

# dropping the columns now add them later again :(, I don't know it better
df_ETF_Share_ts <- df_ETF_Share_ts[ , !names(df_ETF_Share_ts) %in% c("cs_All_Inv", "cs_ETF_Share_Inv")]

# Creating time series for the overall invested amount
df_All_Inv_ts <- df_All_Inv %>%  complete(Date = seq.Date(min(Date), max(Sys.Date()), by= "day")) %>% # Generates the dates between the Date. Sys.Date for update current date
  fill(cs_All_Inv)

# Creating time series for the invested amount per stock type (ETF/Share)
df_ETF_Share_Inv_ts <- df_ETF_Share_Inv %>% group_by(Type)%>%  complete(Date = seq.Date(min(Date), max(Sys.Date()), by= "day")) %>% # Generates the dates between the Date. Sys.Date for update current date
  fill(Type,cs_ETF_Share_Inv)%>% ungroup()

# Creating time series for the trading costs
df_cs_Trad_cost_ts <- df_cs_Trad_cost %>% complete(Date = seq.Date(min(Date), max(Sys.Date()), by= "day")) %>% # Generates the dates between the Date. Sys.Date for update current date
  fill(cs_Trad_cost)

# join them together
df_ETF_Share_ts <- left_join(df_ETF_Share_ts , df_cs_Trad_cost_ts, by = "Date") #Trading cost
df_ETF_Share_ts <- left_join(df_ETF_Share_ts , df_All_Inv_ts, by = "Date") # overall invested amount
df_ETF_Share_ts <- left_join(df_ETF_Share_ts , df_ETF_Share_Inv_ts, by = c("Date", "Type")) # invested amount per stock type (ETF/Share)


# house cleaning
remove(df_cs_Trad_cost, df_cs_Trad_cost_ts, df_All_Inv_ts, df_All_Inv,df_ETF_Share_Inv_ts, df_ETF_Share_Inv )

# getting names of the xts-objects created by getSymbol for the for-loop below
vec_PF_StockNames <- ls(pattern = paste(vec_EPIC, collapse = "|"))


# Need to change xts-object from getSymbols to into a data.frame to merge it
for(i in 1:length(vec_PF_StockNames)){
   df_temp <- data.frame(Date = as.Date(index(get(vec_PF_StockNames[i]))),coredata(get(vec_PF_StockNames[i])))
   assign(paste(vec_PF_StockNames[i]), df_temp)
}

##### Joining the finance data with our portfolio data



for(i in 1:length(vec_EPIC)){
  k = which (str_detect(vec_PF_StockNames, vec_EPIC[i])) # check which EPIC matches which data.frame name for the historical record in vec_PF_StockNames. NOTE: identical EPIC's at different Stock exchange, will cause a problem here
  df_temp <- left_join(df_ETF_Share_ts %>% filter(EPIC == vec_EPIC[i]), get(vec_PF_StockNames[k])[ , c("Date",paste(vec_PF_StockNames[k],".Close",sep = ""))], by = "Date", )
  
  colnames(df_temp)[ncol(df_temp)] <- "Value" # ncol() give us the number of columns in df_temp and coincidentally that number also relates to the last column, which we want to rename. 

# Creating the combined stock timeserie with all stock values in one column
  if(exists("df_temp_rbind")){
    df_temp_rbind <- rbind(df_temp_rbind, df_temp)
  }else{
    df_temp_rbind <- df_temp
  }
}

df_ETF_Share_ts <- df_temp_rbind

rm(df_temp, df_temp_rbind)

```

```{r cleaning the Stock timeseries and calculating the Stock values, include=FALSE }

# removing weekends which are 1 and 7 in wday()-function
df_ETF_Share_ts <-  df_ETF_Share_ts[!(wday(df_ETF_Share_ts$Date) %in% c(1,7)),]
#Note: there will still be NA on Bank holidays, I guess. But they differ from country to country

#### removing other NA
## e.g. bank holidays
df_ETF_Share_ts <- df_ETF_Share_ts[!is.na(df_ETF_Share_ts$Value),]
## Setting a Volume of 0 (When stock has been completely sold) to NA and removing it. To keep the dataset as small as possible
df_ETF_Share_ts$cs_Volume[which(df_ETF_Share_ts$cs_Volume==0)] <- NA
df_ETF_Share_ts <- df_ETF_Share_ts[!is.na(df_ETF_Share_ts$cs_Volume),]


# Calculating the overall Stock value in Portfolio
df_ETF_Share_ts$PF_Stock_Value <- round((df_ETF_Share_ts$cs_Volume * df_ETF_Share_ts$Value)/100,2) # to have it in GBP


# VUSA.L is displayed in GBP and not pence
if(any(levels(df_ETF_Share_ts$EPIC) == "VUSA")){
df_ETF_Share_ts$PF_Stock_Value[df_ETF_Share_ts$EPIC == "VUSA"]<- df_ETF_Share_ts$PF_Stock_Value[df_ETF_Share_ts$EPIC == "VUSA"]*100
}else{
  
}

```


```{r more calculation, include=FALSE}
# relative gains and losses in percent
df_ETF_Share_ts$rel_Gain_Loss <- round(((df_ETF_Share_ts$PF_Stock_Value-df_ETF_Share_ts$cs_Stock_Inv)/df_ETF_Share_ts$cs_Stock_Inv)*100,2)
# relative gains and losses in percent overall Share and ETF


# sum stock values by type and date

df_ETF_Share_ts$RGL_EFTShare <-round(((left_join(df_ETF_Share_ts,aggregate(df_ETF_Share_ts["PF_Stock_Value"], by=df_ETF_Share_ts[c("Date", "Type")],sum), by= c("Date", "Type"))$PF_Stock_Value.y - df_ETF_Share_ts$cs_ETF_Share_Inv)/df_ETF_Share_ts$cs_ETF_Share_Inv)*100,2)  


# Here we are calculating the relative overall gains & losses base on the over investment in the stock type ETF & Share. 1) In the left_join() we are calculating the daily sum of the over ETF and Share values in our portfolio. 2) And then subtract the type based in investment from the daily sum. 3) turn it in percentage  
df_cs_All_value <- data.frame(Date = df_ETF_Share_ts$Date, cs_All_Value = df_ETF_Share_ts$PF_Stock_Value)
df_cs_All_value <- aggregate(df_cs_All_value["cs_All_Value"], by=df_cs_All_value["Date"],sum) # We'll need that step, in case more than one trade action was made per day
df_ETF_Share_ts <- left_join(df_ETF_Share_ts , df_cs_All_value, by = "Date") #Trading cost
rm(df_cs_All_value)


```

# Portfolio distribution



```{r Figure 1. Current distributions, echo=FALSE, message=FALSE, warning=FALSE}

# Overall Market & Sector distribution
subset(df_ETF_Share_ts, Date %in% max(df_ETF_Share_ts$Date))%>%
  group_by(Market, Sector) %>% 
  summarise(across(c(PF_Stock_Value), list(sum = sum))) %>% 
  droplevels.data.frame() %>% 
  PieDonut(., aes(Market, Sector, count=PF_Stock_Value_sum),
         title = "Portfolio Market & Sector distribution",
         color = "black",
                  ratioByGroup = FALSE)


```

Figure 1. A Pie chart showing the portfolio's distribution in the different markets (inner ring) and sectors (outer ring). The *"Mixed"* are EFT's that do not focus on a specific sector (e.g. MSCI world).



```{r Figure 2. Portfolio Sector & Stock distribution, echo=FALSE, message=FALSE, warning=FALSE}
# Overall Sector & Stock distribution
subset(df_ETF_Share_ts, Date %in% max(df_ETF_Share_ts$Date))%>%
  select(Sector, EPIC, PF_Stock_Value) %>% 
  droplevels.data.frame() %>% 
  PieDonut(., aes(Sector, EPIC, count=PF_Stock_Value),
         title = "Portfolio Sector & Stock distribution",
         color = "black",
                  ratioByGroup = FALSE)
```

Figure 2. A Pie chart showing the portfolio's distribution in the different sectors (inner ring) and the respective stocks in the portfolio (outer ring). The *"Mixed"* are EFT's that do not focus on a specific sector (e.g. MSCI world).



```{r Figure 3. ETF Market & Sector distribution, echo=FALSE, message=FALSE, warning=FALSE}
## ETF
subset(df_ETF_Share_ts, Date %in% max(df_ETF_Share_ts$Date))%>%
  filter(Type =="ETF") %>% 
  group_by(Market, Sector) %>% 
  droplevels.data.frame() %>% 
  summarise(across(c(PF_Stock_Value), list(sum = sum))) %>% 
  as.data.frame() %>% 
  droplevels(.$Sector) %>% 
  PieDonut(., aes(Market, Sector, count=PF_Stock_Value_sum),
         title = "ETF Market & Sector distribution",
         color = "black",
                  ratioByGroup = FALSE)
```

Figure 3. A Pie chart showing the portfolio's EFT distribution in the different markets (inner ring) and the respective sector distribution (outer ring). The *"Mixed"* are EFT's that do not focus on a specific sector (e.g. MSCI world). 



```{r Figure 4. Share & Sector distribution, echo=FALSE, message=FALSE, warning=FALSE}
## Share
subset(df_ETF_Share_ts, Date %in% max(df_ETF_Share_ts$Date))%>%
  filter(Type =="Share") %>% 
  group_by(Sector, EPIC) %>% 
  summarise(across(c(PF_Stock_Value), list(sum = sum))) %>% 
  droplevels.data.frame %>% 
  PieDonut(., aes(Sector,EPIC, count=PF_Stock_Value_sum),
         title = "Share & Sector distribution",
         color = "black",
                  ratioByGroup = FALSE)
```

Figure 4. A Pie chart showing the portfolio's share distribution in the different sectors (inner ring) and the respective stocks (outer ring).



```{r Overall Portfolio preformance, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

ggplot(df_ETF_Share_ts,aes(Date,PF_Stock_Value, fill=EPIC))+geom_area(alpha=0.9, colour="black")+
  ylab("Overall portfolio value in GPB")+xlab("")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_x_date(date_labels = "%b-%Y", date_breaks = "month")

ggplot(df_ETF_Share_ts)+
  geom_area(aes(Date,PF_Stock_Value, fill=EPIC),alpha=0.5)+ 
  geom_line(aes(Date,cs_Stock_Inv,colour="Investment"),linetype="dashed")+
  ylab("Overall portfolio value in GPB")+xlab("")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_x_date(date_labels = "%b-%Y", date_breaks = "month")+facet_wrap(~EPIC, ncol = 2)+
  scale_color_manual(name = "", values = c("Investment" = "black")) # colour needs to be aes()


```

# Portfolio performance




```{r Figure 5.  Portfolio growth [GBP], echo=FALSE, message=FALSE, warning=FALSE}
df_ETF_Share_ts %>% 
  select(Date,cs_All_Inv,cs_All_Value,cs_Trad_cost) %>% 
  distinct() %>% 
  ggplot(.,aes(Date,cs_All_Inv))+
  geom_area(aes(x=Date, y=ifelse(cs_All_Value>cs_All_Inv, cs_All_Value,cs_All_Inv)), fill="darkgreen", alpha = 0.7)+
  geom_area(aes(x=Date, y=cs_All_Inv), fill="darkred", alpha = 0.7)+
  geom_area(aes(x=Date, y=ifelse(cs_All_Value<cs_All_Inv, cs_All_Value,cs_All_Inv)), fill="grey") +
  geom_line(aes(Date,cs_Trad_cost,colour="Trading costs"),linetype="dashed")+
  ylab("Portfolio growth [GBP]")+xlab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_x_date(date_labels = "%b-%Y", date_breaks = "month")+
  scale_color_manual(name = "", values = c("Trading costs" = "black"))+
  ggtitle("Portfolio performance over its lifetime")

```

Figure 5. Shows the Portfolio performance over its lifetime with the growth (green) and the shrinking (red) of the investment (grey). The trading costs are represented by the black dashed line.



```{r Figure 6. Relative gain and losses of stocks in the Portfolio, echo=FALSE}

ggplot(df_ETF_Share_ts)+
  geom_area(aes(x=Date, y=ifelse(rel_Gain_Loss<0, rel_Gain_Loss,0)), fill="darkred", alpha = 0.7) +
  geom_area(aes(x=Date, y=ifelse(rel_Gain_Loss>0, rel_Gain_Loss, 0)), fill="darkgreen", alpha = 0.7)+
  geom_hline(yintercept=0, linetype="dashed")+
  ylab("Gain and Losses [%]")+xlab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_x_date(date_labels = "%b-%Y", date_breaks = "month")+facet_wrap(~EPIC, ncol = 2)+
  ggtitle("All relative gains and losses of stocks")




```

Figure 6. Shows the all relative gains (green) and losses (red) of the stocks inside the portfolio based on their the daily closing values. The shown time period is based on the buy and sold actions surrounding the respective stock. If a stock is completely sold, it time series stop. The relative gains and losses are based on the overall investment into the respective stock.



```{r Figure 7. Trading costs, echo=FALSE}

ggplot(df_ETF_Share_ts,aes(Date,cs_Trad_cost))+
  geom_line()+
  ylab("Lifetime Trading cost [GPB]")+xlab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_x_date(date_labels = "%b-%Y", date_breaks = "month")+
  ggtitle("Lifetime trading costs of the portfolio")
```

Figure 7. Shows the trading costs over the lifetime of the portfolio.



```{r Figure 8. Individual ETF performance gains & losses, echo=FALSE}

ggplot(subset(df_ETF_Share_ts, Type %in% "ETF"))+
  geom_area(aes(x=Date, y=ifelse(rel_Gain_Loss<0, rel_Gain_Loss,0)), fill="darkred", alpha = 0.7) +
  geom_area(aes(x=Date, y=ifelse(rel_Gain_Loss>0, rel_Gain_Loss, 0)), fill="darkgreen", alpha = 0.7)+
  geom_hline(yintercept=0, linetype="dashed")+
  ylab("Gain and Losses [%]")+xlab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_x_date(date_labels = "%b-%Y", date_breaks = "month")+facet_wrap(~EPIC, ncol = 2)+
  ggtitle("Relative gains and losses of the individual ETF stocks")


```

Figure 8. Shows the relative gains (green) and losses (red) of the ETF stocks inside the portfolio based on their the daily closing values. The shown time period is based on the buy and sold actions surrounding the respective stock. If a stock is completely sold, it time series stop. The relative gains and losses are based on the overall investment into the respective stock.



```{r Figure 9. ETF performance [GBP], echo=FALSE}
ggplot(subset(df_ETF_Share_ts, Type %in% "ETF"))+
  geom_area(aes(x=Date, y=ifelse(PF_Stock_Value>cs_Stock_Inv, PF_Stock_Value,cs_Stock_Inv)), fill="darkgreen", alpha = 0.7)+
  geom_area(aes(x=Date, y=cs_Stock_Inv), fill="darkred", alpha = 0.7)+
  geom_area(aes(x=Date, y=ifelse(PF_Stock_Value<cs_Stock_Inv, PF_Stock_Value,cs_Stock_Inv)), fill="grey") +
  ylab("ETF performance [GBP]")+xlab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_x_date(date_labels = "%b-%Y", date_breaks = "month")+facet_wrap(~EPIC, ncol = 2)+
  ggtitle("Performance of the individual ETF stocks")
```

Figure 9. Shows the individual ETF stock performance over their lifetime in the portfolio with the growth (green) and the shrinking (red) of the investment (grey). 



```{r Figure 10. Share performance gains & losses, echo=FALSE}

ggplot(subset(df_ETF_Share_ts, Type %in% "Share"))+
  geom_area(aes(x=Date, y=ifelse(rel_Gain_Loss<0, rel_Gain_Loss,0)), fill="darkred", alpha = 0.7) +
  geom_area(aes(x=Date, y=ifelse(rel_Gain_Loss>0, rel_Gain_Loss, 0)), fill="darkgreen", alpha = 0.7)+
  geom_hline(yintercept=0, linetype="dashed")+
  ylab("Gains and Losses [%]")+xlab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_x_date(date_labels = "%b-%Y", date_breaks = "month")+facet_wrap(~EPIC, ncol = 2)+
  ggtitle("Relative gains and losses of the individual share stocks")

```

Figure 10. Shows the relative gains (green) and losses (red) of the share stocks inside the portfolio based on their the daily closing values. The shown time period is based on the buy and sold actions surrounding the respective stock. If a stock is completely sold, it time series stop. The relative gains and losses are based on the overall investment into the respective stock.



```{r Figure 11. Share performance [GBP], echo=FALSE}
ggplot(subset(df_ETF_Share_ts, Type %in% "Share"))+
  geom_area(aes(x=Date, y=ifelse(PF_Stock_Value>cs_Stock_Inv, PF_Stock_Value,cs_Stock_Inv)), fill="darkgreen", alpha = 0.7)+
  geom_area(aes(x=Date, y=cs_Stock_Inv), fill="darkred", alpha = 0.7)+
  geom_area(aes(x=Date, y=ifelse(PF_Stock_Value<cs_Stock_Inv, PF_Stock_Value,cs_Stock_Inv)), fill="grey") +
  ylab("Share preformance [GBP]")+xlab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_x_date(date_labels = "%b-%Y", date_breaks = "month")+facet_wrap(~EPIC, ncol = 2)+
  ggtitle("Performance of the individual share stocks")
```

Figure 11. Shows the Individual share stock performance over their lifetime in the portfolio with the growth (green) and the shrinking (red) of the investment (grey).



```{r Figure 12. Share vs ETF, echo=FALSE}

ggplot(df_ETF_Share_ts)+
  geom_area(aes(x=Date, y=ifelse(RGL_EFTShare<0, RGL_EFTShare,0)), fill="darkred", alpha = 0.7, position = "identity") +
  geom_area(aes(x=Date, y=ifelse(RGL_EFTShare>0, RGL_EFTShare, 0)), fill="darkgreen", alpha = 0.7,position = "identity")+
  geom_hline(yintercept=0, linetype="dashed")+
  facet_wrap(~Type)+
  ylab("Gain and Losses [%]")+xlab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_x_date(date_labels = "%b-%Y", date_breaks = "month")+
  ggtitle("Gains and losses ETF versus Share")


```
Figure 12. Shows the relative gains (green) and losses (red) between ETF and share stocks in the portfolio based on their the daily closing values. The shown time period is based on the buy and sold actions surrounding the respective stock. If a stock is completely sold, it time series stop. The relative gains and losses are based on the overall investment into the respective stock.



# Dividend assessment


```{r Calculating the dividend per stock, message=FALSE, warning=FALSE, include=FALSE}

for(i in 1:length(vec_EPIC)){
  k = which (str_detect(vec_PF_StockNames, vec_EPIC[i])) # check which EPIC matches which data.frame name for the historical record in vec_PF_StockNames. NOTE: identical EPIC's at different Stock exchange, will cause a problem here
  df_temp <- left_join(df_PF_div %>% filter(EPIC == vec_EPIC[i]), get(vec_PF_StockNames[k])[ , c("Date",paste(vec_PF_StockNames[k],".Close",sep = ""))], by = c("XD_Date"="Date"), )
  
  colnames(df_temp)[ncol(df_temp)] <- "Value" # ncol() give us the number of columns in df_temp and coincidentally that number also relates to the last column, which we want to rename. 

# Creating the combined stock timeserie with all stock values in one column
  if(exists("df_temp_rbind")){
    df_temp_rbind <- rbind(df_temp_rbind, df_temp)
  }else{
    df_temp_rbind <- df_temp
  }
}

df_PF_div <- df_temp_rbind

rm(df_temp, df_temp_rbind)


# Converting GBX to GBP
df_PF_div$PF_Stock_Value <- round((df_PF_div$Shares_held_XD_date * df_PF_div$Value)/100,2) # to have it in GBP

# VUSA.L is displayed in GBP and not pence
df_PF_div$PF_Stock_Value[df_PF_div$EPIC == "VUSA"]<- df_PF_div$PF_Stock_Value[df_PF_div$EPIC == "VUSA"]*100

#Calculating the dividend yield

df_PF_div$Div_yield <- round((df_PF_div$Amount_Payable/df_PF_div$PF_Stock_Value)*100,2)

```


```{r Figure 13. Paid out dividends per year, echo=FALSE}

df_PF_div %>% 
  group_by(Year = format(Issue_Date, format="%Y")) %>% 
ggplot(., aes(x = Year,y = Amount_Payable))+geom_bar(stat = "identity", fill="darkgreen", alpha=0.7)+
  ylab("Dividend paid out [GPB]")+xlab("")+
  ggtitle("Dividend paid out per year")

```

Figure 13. 



```{r Figure 14. Dividend payout per year, echo=FALSE, message=FALSE, warning=FALSE}


df_PF_div %>% 
  group_by(EPIC,Year = format(Issue_Date, format="%Y")) %>%
  summarise(Sum = sum(Amount_Payable), Sum_yield = sum(Div_yield), Payouts = n() ) %>% 
ggplot(., aes(x = Year ,y = Sum, fill=EPIC ))+
  geom_col( colour = "Black", position = position_dodge(), alpha=0.7)+
  geom_text(aes(label=paste(Sum_yield,"%",sep=""), group = EPIC),vjust=-.6, position = position_dodge(width=0.9), size=2.5)+
  geom_text(aes(label=paste(Payouts), group = EPIC),vjust=1.2,colour="Darkred", position = position_dodge(width=0.9), size=3)+
  ylab("Dividend paid out [GPB]")+xlab("")
```

Figure 14.

```{r Figure 15., echo=FALSE, message=FALSE, warning=FALSE}
# Life time paid out
df_PF_div %>% 
  group_by(EPIC) %>%
  summarise(sum = sum(Amount_Payable)) %>% 
ggplot(., aes(x = EPIC ,y = sum))+geom_bar(stat = "identity", fill="darkgreen", alpha=0.7)+
  ylab("Dividend paid out [GPB]")+xlab("")

```

Figure 15.

## Dividends vs Trading costs

```{r eval=FALSE, include=FALSE}
df_ETF_Share_ts %>% 
  group_by(Year = format(Date, format="%Y")) %>% 
  
```

